### 2015-07-14 tensor/draft/draft-tensor-op.tex
[draft] check section on exp. inputs

% \documentclass[prd,preprintnumbers,nofootinbib,twocolumn]{revtex4-1}
\documentclass[twocolumn,epjc3]{svjour3}

\RequirePackage[T1]{fontenc}
\RequirePackage{mathptmx}      % use Times fonts if available on your TeX system
\RequirePackage{flushend}
\RequirePackage[numbers,sort&compress]{natbib}
\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue,linkcolor=blue]{hyperref}

\journalname{Eur. Phys. J. C}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{multirow}
\usepackage{xcolor}

\usepackage{textcomp}

\allowdisplaybreaks[1]

%%
%% Setup
%%
% Use the section number in equation tags
\renewcommand{\theequation}{\thesection.\arabic{equation}}
\numberwithin{equation}{section}
\setlength{\parindent}{15pt}

\lstset{
    basicstyle=\ttfamily,
    backgroundcolor=\color{lightgray}
}

%--------+---------+---------+---------+---------+---------+---------+---------+
%
%    Macros
%
%--------+---------+---------+---------+---------+---------+---------+---------+

% references to ...
\def \refeq#1{(\ref{#1})}
\def \refsec#1{Section \ref{#1}}
\def \refapp#1{\ref{#1}}
\def \reffig#1{Figure \ref{#1}}
\def \reftab#1{Table \ref{#1}}

\def \nn{\nonumber\\}
\def \dis{\displaystyle}

%% tables
\newcommand\tabvsptop{\rule{0pt}{2.6ex}}
\newcommand\tabvspbot{\rule[-1.5ex]{0pt}{0pt}}

%%
%% Comments
%%
\newcommand{\eos}[1]{{\color{red}{#1}}}   % things to be checked and replaced with EOS results
\newcommand{\fred}[1]{{\color{green}{#1}}}
\newcommand{\todo}[1]{{\color{magenta}{ {\bf ToDo: }{#1}}}}
\newcommand{\checked}[1]{{\color{brown}{ {\bf Checked: }{#1}}}}

%% ----------------------------
%%  Physics, Math, Convenience
%% ----------------------------

% Physics
\newcommand{\wilson}[2][{}]{\mathcal{C}_{#2}^{\mathrm{#1}}}
\newcommand{\op}[1]{\mathcal{O}_{#1}}
\newcommand{\bra}[1]{\big\langle{#1}\big\vert}
\newcommand{\ket}[1]{\big\vert{#1}\big\rangle}
\newcommand{\order}[1]{\mathcal{O}\left({#1}\right)}

% Convenience
\renewcommand{\[}{\big[}
\renewcommand{\]}{\big]}
\renewcommand{\(}{\big(}
\renewcommand{\)}{\big)}
\newcommand{\dd}{{\rm d}}
\newcommand{\para}{\parallel}
\newcommand{\eps}{\varepsilon}
\newcommand{\GeV}{\ensuremath{\mathrm{GeV}}}
%\newcommand{\MeV}{{\si{\MeV}}
\newcommand{\msbar}{\ensuremath{\overline{\rm MS}}}

%% Physics
\newcommand{\alphas}{\alpha_\mathrm{s}}
\newcommand{\alphae}{\alpha_\mathrm{e}}
\newcommand{\gfermi}{G_\mathrm{F}}
\newcommand{\amp}[1]{\mathcal{A}\left({#1}\right)}

%--------------
% math symbols
%--------------

\renewcommand{\Im}[1]{{\rm Im}\left\lbrace{#1}\right\rbrace}
\renewcommand{\Re}[1]{{\rm Re}\left\lbrace{#1}\right\rbrace}

\def \ket#1{ \left| #1 \right\rangle}
\def \brac#1{\left\langle #1 \right|}
\def \averA#1{\left\langle #1 \right\rangle}
\def \aver#1{\langle #1 \rangle}
\def \abs#1{ \left| #1 \right| }
\def \order#1{ {\cal O} \left( #1 \right) }
\def \dslash#1{ #1\!\!\!/}
\def \re{\textrm{Re}}
\def \im{\textrm{Im}}
\def \eps{\varepsilon}
\def \LogGamma{\textrm{LogGamma}}
\def \Amoroso{\textrm{Amoroso}}
\def \vecth{\vec{\theta}}
\def \vecnu{\vec{\nu}}

%-------------
% caligraphic
%-------------

\def \cA{{\cal A}}
\def \cC{{\cal C}}
\def \cF{{\cal F}}
\def \cL{{\cal L}}

%
% effective Lagrangian, operators
%

\def \cLdB1{{{\cal L}_{\Delta B = 1}^{\rm EW}}} % Delta_B = 1 effective Lagrangian
\def \BR{{\cal B}}                               % branching ratio

\def \Op{{\cal O}}
\def \OpQ{{Q}}
\def \OpF{{O}}         % current operator of full theory (Delta_B = 1)
\def \OpE{{\cal O}}    % current operator of effective theory

\def \One{\leavevmode\hbox{\small1\kern-3.6pt\normalsize1}} % unit matrix

%-------------
% differential
%-------------

\newcommand\rmdx[1]{\mbox{d} \, #1 \,}

%--------------------------------
% couplings, scales, observables
%--------------------------------

\def \gS{g_s}              % strong coupling
\def \alS{\alpha_s}        % strong coupling
\def \alE{\alpha_e}        % electro-magnetic coupling
\def \GF{G_F}              % Fermi coupling
\def \LamConf{{\Lambda_{\rm QCD}}}    %  confinement scale
\def \LamConfS{{\Lambda^2_{\rm QCD}}} % (confinement scale)^2

\def \Gaml{{\Gamma_l}}
\def \Game{{\Gamma_e}}
\def \Gammu{{\Gamma_\mu}}

\def \BRl{{{\cal B}_l}}
\def \BRe{{{\cal B}_e}}
\def \BRmu{{{\cal B}_\mu}}

\def \BRll#1#2#3{{\BR(#1) |_{[#2] #3}}}

\def \AFBl{{A_{\rm FB}^l}}
\def \AFBmu{{A_{\rm FB}^\mu}}

\def \FHl{{F_H^l}}
\def \FHe{{F_H^e}}
\def \FHmu{{F_H^\mu}}

%--------
% angles
%--------

\def \thl {{\theta_\ell}}
\def \thK {{\theta_{K}}}
\def \barthl {{\bar{\theta}_l}}
\def \barthK {{\bar{\theta}_{K^*}}}

%--------
% decays
%--------

%
% B -> X + M
%
\def \BtoXM  {\bar{B} \to XM}
%
% b -> scc, B -> J/psi + ...
%
\def \cc{{\bar{c}c}}    % ccbar
\def \bscc   {b \to s + (c\bar{c})}

\def \BJpsiX {\bar{B} \to J/\psi X}
\def \BJpsiXs{\bar{B} \to J/\psi X_s}
\def \BJpsiK {\bar{B} \to J/\psi K}
%
% b -> clv
%
\def \bclv {b\to c l \bar\nu_l}
\def \BXclv  {\bar{B} \to X_c l \bar\nu_l}
%
% b -> s gamma
%
\def \bsgamma {b\to s \gamma}
\def \BXsgamma {\bar{B} \to X_s \gamma}
\def \BtoKastgamma{{ \bar{B} \to K^{\ast} \gamma}}
%
% bs -> ll, B_q -> ll
%
\def \Bstoll{{\bar{B}_s \to \bar{l}l}}
\def \Bstomm{{\bar{B}_s \to \bar{\mu}\mu}}
\def \Bqtomm{{\bar{B}_{d,s} \to \bar{\mu}\mu}}
%
% b -> sll
%
\def \bsll {b\to s \bar{l}l}
\def \BXsll{\bar{B} \to X_s \bar{l}l}

\def \BtoKast{{ \bar{B} \to K^{\ast}}}
\def \B0toK0ast{{ \bar{B}^0 \to K^{\ast 0}}}
%
% excl B->Kll
%
\def \BtoKll{{\bar{B} \to K \bar{l}l}}
\def \BtoKee{{\bar{B} \to K \bar{e}e}}
\def \BtoKmumu{{\bar{B} \to K \bar{\mu}\mu}}
\def \BptoKpll{{B^+ \to K^+ \bar{l}l}}
\def \BmtoKmll{{B^- \to K^- \bar{l}l}}
\def \BneutralKll{{\bar{B}^0 \to K^0 \bar{l}l}}
%%%\def \B0toK0ll{{\bar{B}^0 \to K^0 \bar{l}l}}
\def \BtoKastll{{\bar{B} \to K^\ast \bar{l}l}}
\def \BtoKastmumu{{\bar{B} \to K^\ast \bar{\mu} \mu}}

\def \BmtoKmmumu{{ B^- \to K^- \bar{\mu} \mu}}
\def \B0toK0mumu{{ \bar{B}^0 \to K^0 \bar{\mu} \mu}}
%
% excl B-> K^* (-> K pi) ll
%
\def \BtoKKpill{{ B \to K^\ast (\to K \pi) \bar{l}l}}
\def \BtoKKpimm{{ B \to K^\ast (\to K \pi) \bar{\mu}\mu}}
\def \barB0toKKpill{{ \bar{B}^0 \to \bar{K}^{\ast 0} (\to K^- \pi^+) \bar{l}l}}
\def \B0toKKpill{{ B^0 \to K^{\ast 0} (\to K^+ \pi^-) \bar{l}l}}
\def \BtoKastKpill{{ (\bar{B}^0, B^0)\to (K^{\ast 0},\bar{K}^{\ast 0}) (\to K^\mp \pi^\pm) \bar{l}l}}

%
% Scenario Names
%
\def \SMnu{SM($\nu$-only)}
\def \SM{SM}
\def \SMp{SM+SM$'$}
\def \SMpNine{SM+SM$'$$(9)$}

%
% useful variable for multiple same-size plots
%
\newlength{\relwidth}
%--------+---------+---------+---------+---------+---------+---------+---------+
%
%
%
%--------+---------+---------+---------+---------+---------+---------+---------+

\begin{document}

% better line breaks, http://tex.stackexchange.com/questions/32062
\sloppy

\title{Constraints on tensor and scalar couplings from
  $B\to K\bar\mu\mu$ and $B_s\to \bar\mu\mu$
}
\author{Frederik Beaujean\thanksref{e1,addr1}
  \and
  Christoph Bobeth\thanksref{e2,addr2}
  \and
  Stephan Jahn\thanksref{e3,addr3,addr4}
}
\institute{
  C2PAP, Excellence Cluster Universe,
  Ludwig-Maximilians-Universit\"at M\"unchen,
  Garching, Germany\label{addr1}
  \and
  Institute for Advanced Study,
  Technische Universit\"at M\"unchen,
  Garching, Germany\label{addr2}
  \and
  Excellence Cluster Universe,
  Technische Universit\"at M\"unchen,
  Garching, Germany\label{addr3}
  \and
  \emph{Present Address:}
  Max Planck institute for physics, Munich, Germany\label{addr4}
}

\thankstext[$\star$]{t1}{\todo{EOS-2015-??}, FLAVOUR(267104)-ERC-107}
\thankstext{e1}{e-mail: frederik.beaujean@lmu.de}
\thankstext{e2}{e-mail: christoph.bobeth@ph.tum.de}
\thankstext{e3}{e-mail: sjahn@mpp.mpg.de}

\date{\today}
\maketitle

\begin{abstract}
  \todo{check}
  The angular distribution of $B\to K\bar\ell\ell$ ($\ell =
  e,\,\mu,\,\tau$) provides a lepton-forward-backward asymmetry,
  $A_{\rm FB}^\ell$, and a flat term, $F_H^\ell$, that are strongly
  suppressed in the standard model. In both, nonstandard tensor and
  scalar contributions are enhanced with respect to vector ones. We
  use latest experimental results for $\ell = \mu$ in combination with
  $B_s\to \bar\mu\mu$ to derive stronger model-independent constraints
  on tensor and scalar effective couplings and determine their impact
  on not yet measured angular observables in $B\to K^*\bar\mu\mu$,
  which feature the same suppression of vector couplings.  Most
  notably, the measurement of $F_H^\mu$ provides constraints on
  (pseudo-)scalar couplings and their chirality-flipped counterparts
  that are complementary to the ones from the branching ratio $B_s\to
  \bar\mu\mu$, which allows to put for the first time simultaneously
  upper bounds on all four couplings. \fred{Try to shorten, add
    sentence ``independent of deviations in vector-coupling dominated
    observables''. There could be a bit of summary of outcome here, too}
  \todo{check captions of figures and tables}
\end{abstract}

%
%
%
%--------+---------+---------+---------+---------+---------+---------+---------+
\section{
 \checked{ Introduction}
}

With the analysis of the data collected by the LHCb Collaboration
during phase I at the Large Hadron Collider (LHC), we now have access
to rather large samples of rare $B$ meson decays with branching ratios
below $10^{-5}$. As a consequence, angular analyses of three- and
four-body final states can be used to measure a larger number of
observables than previously possible at the B factories BaBar and
Belle. In this work we focus on rare $B$ decays driven at the parton
level by the flavor-changing neutral-current (FCNC) transition $b\to s
\bar\ell\ell$ that constitutes a valuable probe of the standard model
(SM) and provides constraints on its extensions.

The angular distribution of $B\to K \bar\ell\ell$---normalized to the
width $\Gamma_\ell$---in the angle $\theta_\ell$ between $B$ and
$\ell^-$ as measured in the dilepton rest frame is
\begin{align}
  \label{eq:ang-distr}
  \frac{1}{\Gamma_\ell} \frac{\mbox{d}\Gamma_{\ell}}{\mbox{d}\!\cos\theta_\ell} &
  = \frac{3}{4} (1 - F_H^\ell) \sin^2\!\theta_\ell
  + \frac{1}{2} F_H^\ell
  + A_{\rm FB}^\ell \cos\theta_\ell \,.
\end{align}
LHCb recently analyzed their 3 fb$^{-1}$ data set and measured the angular
distribution of the mode $B^+\to K^+ \bar\mu\mu$, i.e. $\ell =
\mu$~\cite{Aaij:2014tfa}, with unprecedented precision to provide the
lepton-forward-backward asymmetry $A_{\rm FB}^\mu$ and the flat term $F_H^\mu$
in CP-averaged form and integrated over several bins in the dilepton invariant
mass $q^2$. Similarly, the CP-averaged branching ratios, ${\cal B}_\mu =
\tau_{B} \Gamma_\mu$,~\cite{Aaij:2014pli} and the rate CP asymmetry $A_{\rm
  CP}^\mu$ \cite{Aaij:2014bsa} are also available from 3 fb$^{-1}$.

Both observables, $F_H^\ell$ and $A_{\rm FB}^\ell$, exhibit strong suppression
factors for vector and dipole couplings present in the SM, thereby enhancing
their sensitivity to tensor and scalar couplings \cite{Bobeth:2007dw,
  Bobeth:2012vn}. A similar enhancement of scalar couplings compared to
helicity-suppressed vector couplings of the SM is well-known from $B_s\to
\bar\mu\mu$. Unfortunately the limited data set of $B\to K^*(\to K\pi)
\bar\ell\ell$ from LHCb \cite{Aaij:2013qta} did not allow to perform a full
angular analysis without the assumption of vanishing scalar and tensor couplings
in this decay mode. In the future with more data or special-purpose analysis
techniques like the method of moments~\cite{beaujean:2015mom}, certain angular
observables in $B\to K^* \bar\ell\ell$ will provide additional constraints on
such couplings, as for example $J_{6c}$ \cite{Altmannshofer:2008dz} and the
linear combinations $(J_{1s} - 3 J_{2s})$ and $(J_{1c} + J_{2c})$
\cite{Matias:2012xw, Bobeth:2012vn} as well as the experimental test of the
relations $H_T^{(2)} = H_T^{(3)}$ and $J_7 = 0$~\cite{Bobeth:2012vn} at low
hadronic recoil.

Here we exploit current data from $B^+\to K^+ \bar\mu\mu$ and $B_s\to
\bar\mu\mu$ to derive stronger constraints than before on tensor and scalar
couplings in various model-independent scenarios and study their impact on the
not yet measured sensitive observables in $B\to K^* \bar\ell\ell$. In
\refsec{sec:EFT}, we specify the effective theory of $|\Delta B| = 1$ decays on
which our model-independent fits are based. Within this theory, we discuss the
dependence of observables in $B\to K \bar\ell\ell$ and $B\to K^* \bar\ell\ell$
on the tensor and scalar couplings in \refsec{sec:observables} and specify also
the experimental input used in the fits. The constraints on tensor and scalar
couplings from the data are presented for several model-independent scenarios in
\refsec{sec:results}.  Technical details of the angular observables in $B\to K^*
\bar\ell\ell$, the branching fraction of $B_s\to \bar\mu\mu$, the treatment of
theory uncertainties, and the Monte Carlo methods used are relegated to
appendices.

%
%
%
%--------+---------+---------+---------+---------+---------+---------+---------+
\section{
  \checked{Effective Theory}
  \label{sec:EFT}
}

In the framework of the $|\Delta B| = |\Delta S| = 1$ effective theory
\begin{equation}
  \label{eq:Heff}
\begin{aligned}
  {\cal L}_{\rm eff}  =
   \frac{4 G_F}{\sqrt{2}} \,\frac{\alE}{4 \pi}\, & V_{tb}^{} V_{ts}^\ast
      \, \Big[ C_7 \Op_7 + C_{7'} \Op_{7'}
 \\ & +
       \sum_{\ell = e,\, \mu,\, \tau} \sum_i  \wilson[\ell]{i}(\mu_b)  \Op_i \Big] + \text{h.c.}
\end{aligned}
\end{equation}
the most general dimension-six flavor-changing operators $\Op_i$ mediating $b\to
s \gamma$ and $b\to s \bar\ell\ell$ are classified according to their chiral
structure. There are dipole ($i = 7,7'$ \fred{so there is no $\mu, \tau$ version
  of $\wilson{7}$? I guess it should also depend on $\mu_b$}) and vector ($i =
9,9',10,10'$) operators
\begin{equation}
  \label{eq:SM:ops}
\begin{aligned}
  \Op_{7(7')} &
  = \frac{m_b}{e}\!\[\bar{s} \sigma^{\mu\nu} P_{R(L)} b\] F_{\mu\nu}\,,
\\
  \Op_{9(9')} &
  = \[\bar{s} \gamma_\mu P_{L(R)} b\]\!\[\bar{\ell} \gamma^\mu \ell\]\,,
\\[0.1cm]
  \Op_{10(10')} &
  = \[\bar{s} \gamma_\mu P_{L(R)} b\]\!\[\bar{\ell} \gamma^\mu \gamma_5 \ell\]\,,
\end{aligned}
\end{equation}
further scalar ($i = S,S',P,P'$) operators
\begin{equation}
  \label{eq:scalar:ops}
\begin{aligned}
  \Op_{S(S')} &
  = \[\bar{s} P_{R(L)} b\]\!\[\bar{\ell} \ell\]\,,
\\[0.1cm]
  \Op_{P(P')} &
  = \[\bar{s} P_{R(L)} b\]\!\[\bar{\ell} \gamma_5 \ell\]\,,
\end{aligned}
\end{equation}
and tensor ($i = T,T5$) operators
\begin{equation}
  \label{eq:tensor:ops}
\begin{aligned}
  \Op_{T} &
  = \[\bar{s} \sigma_{\mu\nu} b\]\!\[\bar{\ell} \sigma^{\mu\nu} \ell\]\,,
\\[0.1cm]
  \Op_{T5} &
  = \[\bar{s} \sigma_{\mu\nu} b\]\!\[\bar{\ell} \sigma^{\mu\nu} \gamma_5 \ell\]\,,
\end{aligned}
\end{equation}
where the notation $\Op_{T5} = i/2\, \varepsilon^{\mu\nu\alpha\beta}
\[\bar{s} \sigma_{\mu\nu} b\]\!\[\bar{\ell} \sigma_{\alpha\beta} \ell\]$ is
also used frequently in the literature. The respective short-distance couplings,
the Wilson coefficients $\wilson[\ell]{i}(\mu_b)$ are evaluated at a scale of
the order of the $b$-quark mass $\mu_b \sim m_b$ and can be modified from SM
predictions in the presence of new physics.

The SM values $\wilson[(\ell)]{7,9,10}$ are obtained at next-to-next-to leading order
(NNLO) \cite{Bobeth:2003at, Huber:2005ig} and depend on the fundamental parameters
of the top-quark and $W$-boson masses, as well as on the sine of the weak mixing angle.
Moreover, they are universal for the three lepton flavors $\ell = e,\, \mu,\, \tau$.
All other Wilson coefficients are numerically suppressed or zero:
$\wilson[SM]{7'} = m_s/m_b\, \wilson[SM]{7}$, $\wilson[\ell, SM]{S,S',P,P'} \sim
{\cal O}(m_b^2/m_W^2)$, and $\wilson[\ell, SM]{9',10',T,T5} = 0$.
%Beyond the SM,
%lepton-flavor non-universal effects are possible in principle and have been
%discussed for example recently \cite{Alonso:2014csa, Hiller:2014yaa, Ghosh:2014awa,
%Biswas:2014gga, Hurth:2014vma, Glashow:2014iga, Altmannshofer:2014rta, Hiller:2014abc} in
%connection with the measurement of $R_K = {\cal B}_\mu/{\cal B}_e$ in
%$q^2 \in [1, 6]$ GeV$^2$ \cite{Aaij:2014ora}, which deviates by $2.6\sigma$
%from the SM prediction $R_K^{\rm SM} = 1.00$ \cite{Bobeth:2007dw, Bouchard:2013mia}.
The Wilson coefficients of the four-quark current-current and QCD-penguin
operators as well as of the chromomagnetic dipole operators are set to their
NNLO SM values at $\mu_b = 4.2 \, \mbox{GeV}$ \cite{Bobeth:2003at, Huber:2005ig}.

%
%
%
%--------+---------+---------+---------+---------+---------+---------+---------+
\section{
  \checked{Observables and experimental input}
  \label{sec:observables}
}

The full dependence of $F_H^\ell$ and $A_{\rm FB}^\ell$ on tensor and scalar
couplings has been presented in \cite{Bobeth:2007dw, Bobeth:2012vn}, adopting
the effective theory \refeq{eq:Heff}, i.e. neglecting higher dimensional operators
with $\mbox{dim} > 6$. These results imply that for SM values of the effective
couplings
\begin{align}
  F_H^\ell(q^2) \big|_{\rm SM} & \propto \frac{m_\ell^2}{q^2}\,, &
  A_{\rm FB}^\ell \big|_{\rm SM} & = 0 \,.
\end{align}
Hence, for $\ell = e,\, \mu$ both observables are quasi-null tests.  The flat
term $F_H^\ell(q^2)|_{\rm SM}$ is strongly suppressed away from both \fred{Why
  both? At $q^2_{\rm max}$ it is strongly suppressed according to the formula
  given.} kinematic endpoints in $q^2$, except for $\ell = \tau$,
\cite{Bobeth:2007dw, Bobeth:2011nj, Bouchard:2013eph}.  Non-zero values of
$A_{\rm FB}^\ell|_{\rm SM}$ can be induced by higher order QED corrections,
which will modify the simple $\cos\theta_\ell$ dependence of the angular
distribution~\refeq{eq:ang-distr}, however, currently there is no solid estimate
available for this source of SM background\footnote{The leading QED corrections
  to $B\to X_s\bar\ell\ell$ \cite{Huber:2015sra} might provide an estimate of
  corresponding effects in $B\to K \bar\ell\ell$.}.

This picture does not change in the presence of new physics contributions to
vector and dipole operators $i = 7,7',9,9',10,10'$. On the other hand,
non-vanishing tensor or scalar contributions are enhanced unless the dynamics of
NP implies similar suppression factors, i.e., lepton-Yukawa couplings for
$F_H^\ell$ or $\alpha_e$ in the case of $A_{\rm FB}^\ell$. In particular,
$F_H^\ell$ is very sensitive to tensor couplings (see \refeq{eq:AFB-dependence}
below) and $A_{\rm FB}^\ell$ is sensitive to the interference of tensor and
scalar couplings (see \refeq{eq:FH-dependence} below).

\begin{figure*}
  \setlength{\relwidth}{0.33\textwidth}
  \begin{center}
    \includegraphics[width=\relwidth]{plots/pdf/ct_K_FH1to6}
    \includegraphics[width=\relwidth]{plots/pdf/ct_K_FH15to22}
    \\
    \includegraphics[width=\relwidth]{plots/pdf/ct_K_BR1to6}
    \includegraphics[width=\relwidth]{plots/pdf/ct_K_BR15to22}
    \\
    \includegraphics[width=\relwidth]{plots/pdf/ct_Kstar_BR1to6}
    % \includegraphics[width=\relwidth]{plots/pdf/ct_Kstar_BR14to16}
    \includegraphics[width=\relwidth]{plots/pdf/ct_Kstar_BR16to19}
    \\
    \includegraphics[width=\relwidth]{plots/pdf/ct_Kstar_J_1c_plus_J_2c1to6}
    \includegraphics[width=\relwidth]{plots/pdf/ct_Kstar_J_1c_plus_J_2c15to19}
    \\
    \includegraphics[width=\relwidth]{plots/pdf/ct_Kstar_J_1s_minus_3J_2s1to6}
    \includegraphics[width=\relwidth]{plots/pdf/ct_Kstar_J_1s_minus_3J_2s15to19}
  \end{center}
\caption{ The sensitivity to the tensor coupling
      $\wilson{T}$ of $F_H$ [first row] and ${\cal B}$ [second row] in
      $B^+ \to K^+ \bar\mu\mu$ and ${\cal B}$ [third row], $J_{1c} +
      J_{2c}$ [fourth row], and $J_{1s}-3 J_{2s}$ [fifth row] in $B^0 \to K^{*0}
      \bar\mu\mu$ integrated over low- and high-$q^2$ bins. In the
      last two rows, the observables are normalized by the lifetime of
      the $B$ meson for easier comparison with the branching
      ratios. The bands represent the theory uncertainties at 68\%
      and 95\% probability.  For comparison, two sets of bands are
      shown for the SM, $\wilson[NP]{9} = 0$ (blue), and
      $\wilson[NP]{9} = -1.1$ (red), and an additional band of the
      latest 68\% uncertainty reported by LHCb (grey) if a measurement
      is available.
      }
  \label{fig:FHvsCT}
\end{figure*}

There are some angular observables $J_i$ in $B\to K^*(\to K\pi)\, \bar\ell\ell$
with the same properties; i.e. tensor and scalar contributions are kinematically
enhanced by a factor $\sqrt{q^2}/m_\ell$ over vector ones present in the SM or
their respective interference terms. These are $J_{6c}$ and the two linear
combinations $(J_{1s} - 3 J_{2s})$ and $(J_{1c} + J_{2c})$ with explicit
formulas given in \refapp{app:BKstarellell}. In our fits and predictions we
include all kinematically suppressed terms. But for the purpose of illustration,
we now consider the analytical dependence for vanishing lepton mass. In this
limit,
\begin{align}
  J_{6c} & \propto
    \mbox{Re}\big[(\wilson[]{P} - \wilson[]{P'})\, \wilson[\ast]{T}
                - (\wilson[]{S} - \wilson[]{S'})\, \wilson[\ast]{T5}\big]
\intertext{is sensitive to the interference of tensor and scalar operators,
  complementary to $A_{\rm FB}^\ell$ in $B\to K \bar\ell\ell$}
  \label{eq:AFB-dependence}
  A_{\rm FB}^\ell & \propto
    \mbox{Re}\big[(\wilson[]{P} + \wilson[]{P'})\, \wilson[\ast]{T5}
                + (\wilson[]{S} + \wilson[]{S'})\, \wilson[\ast]{T}\big]
                \Big/ \Gamma_\ell \,,
\end{align}
i.e., with an interchange of tensor contributions $T \leftrightarrow T5$.  We
note also that $J_{6c}$ contributes to the lepton forward-backward asymmetry of
$B\to K^* \bar\ell\ell$ being $\propto (J_{6s} + J_{6c}/2)$. Since it has to
compete with $J_{6s}$ in this observable, a separate measurement of $J_{6s}$ and
$J_{6c}$ is necessary.

Only tensor contributions enter
\begin{align}
  (J_{1s} - 3\, J_{2s}) & \propto
  \left(\ldots |\wilson{T}|^2 + \ldots |\wilson{T5}|^2 \right) \,,
\end{align}
where the dots indicate different kinematic and form-factor dependencies.  But
tensor \emph{and} scalar contributions enter
\begin{equation}
\begin{aligned}
  (J_{1c} + J_{2c}) & \propto
      \ldots \left(|\wilson{T}|^2 + |\wilson{T5}|^2 \right)
\\
  & + \ldots \left(|\wilson{S} - \wilson{S'}|^2 + |\wilson{P} - \wilson{P'}|^2 \right) \,,
\end{aligned}
\end{equation}
which is similar to the dependence of $F_H^\ell$ in $B\to K
\bar\ell\ell$
\begin{equation}
  \label{eq:FH-dependence}
\begin{aligned}
  F_H^\ell  \propto & \Big[
      \ldots \left(|\wilson{T}|^2 + |\wilson{T5}|^2\right)
\\
  & + \ldots \left(|\wilson{S} + \wilson{S'}|^2 + |\wilson{P} + \wilson{P'}|^2 \right)
  \Big]
  \Big/ \Gamma_\ell \,.
\end{aligned}
\end{equation}
Concerning $F_H^\ell$, the involved kinematic factors---see \cite{Bobeth:2007dw,
  Bobeth:2012vn}---are such that tensor and scalar couplings contribute only
constructively/cumulatively, apart from cancellations among $\wilson{S(P)}$ and
$\wilson{S'(P')}$.  Interference terms in the numerator of $F_H^\ell$ of the
form $(\wilson{T} \times \wilson{7,7',9,9'})$ and $(\wilson{P,P'} \times
\wilson{10,10'})$ are suppressed by $m_\ell/\sqrt{q^2}$. They can only become
relevant in case $\wilson{T} \ll \wilson{7,7',9,9'}$ or $\wilson{P,P'}
\ll\wilson{10,10'}$.

On the one hand, the observables $F_H^\ell$ \refeq{eq:FH-dependence} and $A_{\rm
  FB}^\ell$ \refeq{eq:AFB-dependence} are measured in the angular distribution
\refeq{eq:ang-distr} of $B\to K \bar\ell\ell$ normalized to the decay width
$\Gamma_\ell$ such that uncertainties due to form factors can cancel in part
\cite{Bobeth:2007dw, Bobeth:2012vn}. On the other hand, $J_{6c}$, $(J_{1s} - 3\,
J_{2s})$, and $(J_{1c} + J_{2c})$ appear in the unnormalized angular
distribution of $B\to K^* \bar\ell\ell$. ``Optimized'' versions $S_1$, $M_1$,
and $M_2$ for the low-$q^2$ region for which form factors cancel in the limit of
$m_b \to \infty$ have been identified in~\cite{Matias:2012xw}.  For the
high-$q^2$ region, potential normalizations are discussed in
\refapp{app:BKstarellell}, which could serve to form optimized observables for
special scenarios of either vanishing chirality-flipped vector or tensor or
scalar couplings. In the most general case, however, there are no optimized
observables at high $q^2$. Although form factors do not cancel in this case, it
might still be preferable to use normalizations, for example when the overall
normalization of $B\to K^*$ form factors constitutes a major theoretical
uncertainty.

\begin{table*}
\begin{center}
\renewcommand{\arraystretch}{1.4}
\begin{tabular}{lccc}
\hline
Channel & Constraints & Kinematics & Source
\\
\hline \hline
$B_s\to \bar\mu\mu$
& $\int\dd\tau \BR(\tau)$ &  --  & \cite{Aaij:2013aka,Chatrchyan:2013bka,CMS:2014xfa}
\\
\hline
\multirow{2}{*}{${B^+\to K^+ \bar\mu\mu}$}
    & \multirow{2}{*}{$\BR_\mu$}
    & $q^2\in [1,\, 6],\, [14.18,\, 16],\, [> 16]$ GeV$^2$  & \cite{CDF:2012:BKstarll} \\
    &
    &  $\,\,\,\,\, q^2 \in [1.1,\, 6.0],\, [15.0,\, 22.0]$ GeV$^2$  & \cite{Aaij:2014pli} \\
    & $A_{\rm FB}^\mu$
    &  $\,\,\,\,\, q^2 \in [1.1,\, 6.0],\, [15.0,\, 22.0]$ GeV$^2$  & \cite{Aaij:2014tfa, CDF:2012:BKstarll} \\
    & $F_{H}^\mu$
    &  $\,\,\,\,\, q^2 \in [1.1,\, 6.0],\, [15.0,\, 22.0]$ GeV$^2$  & \cite{Aaij:2014tfa} \\
    & $A_{\rm CP}^\mu$
    &  $\,\,\,\,\, q^2 \in [1.1,\, 6.0],\, [15.0,\, 22.0]$ GeV$^2$  & \cite{Aaij:2014bsa}
\\
\hline
    ${B^0\to K^{\ast 0} \bar\mu\mu}$
    & $\BR_\mu$
    & $q^2\in [1,\, 6],\, [14.18,\, 16],\, [> 16]$ GeV$^2$  & \cite{CDF:2012:BKstarll, Aaij:2013iag,Chatrchyan:2013cda} \\
    & $A_{\rm FB}^\mu$
    & $q^2\in [1,\, 6],\, [14.18,\, 16],\, [> 16]$ GeV$^2$  & \cite{CDF:2012:BKstarll, Aaij:2013iag,Chatrchyan:2013cda} \\
    & $A_{\rm CP}^\mu$
    &  $\,\,\,\,\, q^2 \in [1.1,\, 6.0],\, [15.0,\, 22.0]$ GeV$^2$  & \cite{Aaij:2014bsa}
\\
\hline
${B\to K}$ form factors
    & $f_{0,+,T}$   &  $q^2 = 17,\, 20,\, 23$ GeV$^2$  & \cite{Bouchard:2013eph} \\
\hline
\multirow{2}{*}{${B\to K^*}$ {form factors}}
    & $V,\,A_{0,1,2},\, T_{1,2,3}$ &  $q^2 = 0.1,\, 4.1,\, 8.1,\, 12.1$ GeV$^2$  & \cite{Straub:2015ica} \\
    & $V,\,A_{0,1,2},\, T_{1,2,3}$
    & $q^2 \in [11.9,\, 17.8]$ GeV$^2$  & \cite{Horgan:2013hoa, Horgan:2015vla}  \\
\hline
\end{tabular}
\renewcommand{\arraystretch}{1.0}
\caption{\label{tab:observables} List of all observables of the various
  $b\to s \bar\mu\mu$ decays that enter the fits with respective kinematics
  and experiments that provide the measurements. LCSR and lattice results of
  $B\to K^{(*)}$ form factors are used to constrain parameters of their
  $q^2$-dependence.
  For more details see the text and \refapp{app:theory:inputs}.
}
\end{center}
\end{table*}

To illustrate the sensitivity of $F_H^\ell$ to tensor couplings, we compare it
in \reffig{fig:FHvsCT} to the branching ratios of $B\to K^{(*)} \bar\ell\ell$
for {$\ell = \mu$}, integrated over one low-$q^2$ and one high-$q^2$ bin. The
details of the numerical input and the uncertainty propagation can be found in
\refapp{app:theory:inputs} and \refapp{app:mc}.  In light of the hint of new
physics in $\wilson{9}$ from recent global analyses of $b\to s (\gamma,\,
\bar\ell\ell)$ data \cite{Descotes-Genon:2013wba, Altmannshofer:2013foa,
  Beaujean:2013soa, Hurth:2014vma, Altmannshofer:2014rta}, we show predictions
for $\wilson[NP]{9} = -1.1$ in addition to $\wilson[NP]{9} = 0$.

From \reffig{fig:FHvsCT}, the highest sensitivity to tensor couplings of any $B
\to K$ observable is attained by $F_H^\ell$ at high $q^2$ due to a partial
cancellation of form factors \cite{Bobeth:2012vn}. If the experimental
uncertainty could be reduced further, $F_H^\ell$ would give a very strong
constraint on a simultaneous negative shift in $\wilson[NP]{9}$ and
$\wilson[NP]{T}$. The prediction of ${\cal B}(B\to K \bar\ell\ell)$ is
essentially insensitive to $\wilson{T}$ but sensitive to $\wilson{NP}[9]$. A
stronger impact on global fits, however, would required reduced theory
uncertainty.

The observable ${\cal B}(B\to K^* \bar\ell\ell)$ shows moderate dependence on
$\wilson{T}$ at least at low $q^2$ and has some impact on the constraints on
tensor couplings as will be discussed in \refsec{sec:results}. At the moment,
theory and experimental uncertainty are of similar size.

$J_{1c}+J_{2c}$ is sensitive to $\wilson{T}$ in both $q^2$ regimes. At low
$q^2$, it is mildly affected by $\wilson[NP]{9}$, whereas at high $q^2$ it is
unaffected. Regarding $J_{1s}-3J_{2s}$, the situation is reversed: here the
strong dependence on $\wilson{T}$ appears at low $q^2$.  Overall, $F_H^\ell$,
$J_{1c}+J_{2c}$ at large $q^2$, and $J_{1s}-3J_{2s}$ are sensitive to
$\wilson{T}$ and theoretically very clean around $\wilson{T}=0$.

From the available measurements, $F_H^\ell$ at high $q^2$ currently provides the
most stringent constraints on the size of tensor couplings. Moreover, the
dependence on vector couplings is such that $\wilson[NP]{9} < 0$ leads to
stronger constraints on $\wilson{T}$ than $\wilson[NP]{9} \approx
0$.

Important additional constraints on scalar couplings come from the
branching ratio of $B_s \to \bar\mu\mu$ as given in
\refeq{eq:Br-Bsmumu}. It provides the most stringent constraints on
the moduli $|\wilson{S} - \wilson{S'}|$ and $|\wilson{P} -
\wilson{P'}|$ and further depends only on $(\wilson{10} -
\wilson{10'})$. Thus it is complementary to $F_H^{\ell}$ in $B\to K
\bar\ell\ell$; see Eq. \refeq{eq:FH-dependence}.

Eventually we also explore the effect of interference with NP in the vector
couplings $\wilson{9,\,9',\,10,\,10'}$ on the bounds on tensor and scalar
couplings.  For this purpose we include also the branching ratio, the lepton
forward-backward asymmetry, and the rate CP asymmetry of $B\to K^* \bar\mu\mu$
as they provide additional constraints on the real and imaginary parts of
$\wilson{9,\,9',\,10,\,10'}$.

The experimental inputs of all observables entering our fits is listed
in \reftab{tab:observables} together with input for the $B\to K^{(*)}$
form factors. More details on the latter can be found in
\refapp{app:theory:inputs}.

%
%
%
%--------+---------+---------+---------+---------+---------+---------+---------+
\section{
  Fits and constraints
  \label{sec:results}
}

As a matter of the fact there are no discrepancies between the latest measurements
for $\ell = \mu$ (throughout this section) of $F_H^\mu$ and $A_{\rm FB}^\mu$ in
$B^+ \to K^+ \bar\mu\mu$ and
their tiny SM predictions. Thus our main objective is the derivation of stronger
constraints on tensor and scalar couplings, taking advantage of their enhancement
in both observables over vector ones. For this purpose, we will consider several
model-independent scenarios, starting from rather restricted to more general ones,
in order to asses the effect of cancellations due to interference of various
contributions.

We will start with the scenario of only tensor couplings and see that they are
well constrained by $F_H^\mu$ alone. In a second scenario we consider only scalar
couplings in order to investigate the complementarity of $F_H^\mu$ and
$B_s \to \bar\mu\mu$. Here we find that for the first time all scalar couplings
can be bounded simultaneously by the combination of both measurements. Eventually
we will consider as a special scenario the SM
augmented by dimension six operators as an effective theory of new physics below
some high scale $\Lambda_{\rm NP}$, much larger than the typical scale of electroweak
symmetry breaking $v$ in the presence of one scalar doublet under $SU(2)_L$,
as in the SM, i.e. $v \ll \Lambda_{\rm NP}$. For each scenario, we investigate
also interference effects with new physics in vector couplings. Finally, we
conclude this section with postirior predictions of the allowed ranges of the yet
not measured angular observables in $B\to K^*\bar\mu\mu$ and $A_{\Delta\Gamma}$ in
$B_s\to\bar\mu\mu$ given all other experimental constraints.

\todo{Explain special initialisation of Markov chains at fixed point in parameter
space in order to avoid other local modes in subspace of nuisance parameters.
What would be appropriate place to do so?}

%
%
%--------+---------+---------+---------+---------+---------+---------+---------+
\subsection{Tensor couplings \label{sec:tensor}}

\begin{table*}
  \renewcommand{\arraystretch}{1.3}
  \begin{center}
  \begin{tabular}{cccc}
  \hline
  data set
  & only $F_H^\mu$
  & $F_H^\mu$ + other
  & $F_H^\mu$ + other
  \\
  set of couplings
  & $\wilson{T,\,T5}$
  & $\wilson{T,\,T5}$
  & $\wilson{T,\,T5,\, 9,\,10}$
  \\
  \hline
  $\mbox{Re}\,\wilson{T}$
  & $\quad$ $[-0.32,\, 0.16]$ ($[-0.52,\, 0.35]$) $\quad$
  & $\quad$ $[-0.29,\, 0.09]$ ($[-0.43,\, 0.27]$) $\quad$
  & $\quad$ $[-0.23,\, 0.14]$ ($[-0.39,\, 0.30]$) $\quad$
  \\
  $\mbox{Im}\,\wilson{T}$
  & $[-0.25,\, 0.24]$ ($[-0.44,\, 0.44]$)
  & $[-0.19,\, 0.21]$ ($[-0.37,\, 0.35]$)
  & $[-0.17,\, 0.22]$ ($[-0.33,\, 0.37]$)
  \\[0.1cm]
  $\mbox{Re}\,\wilson{T5}$
  & $[-0.25,\, 0.24]$ ($[-0.44,\, 0.44]$)
  & $[-0.19,\, 0.17]$ ($[-0.33,\, 0.33]$)
  & $[-0.18,\, 0.16]$ ($[-0.32,\, 0.32]$)
  \\
  $\mbox{Im}\,\wilson{T5}$
  & $[-0.24,\, 0.25]$ ($[-0.44,\, 0.45]$)
  & $[-0.21,\, 0.19]$ ($[-0.37,\, 0.35]$)
  & $[-0.20,\, 0.18]$ ($[-0.36,\, 0.35]$)
  \\
  \hline
  $|\wilson{T}|$, $|\wilson{T5}|$
  & $[0.13,\, 0.43]$ ($[0.03,\, 0.57]$)
  & $[0.09,\, 0.33]$ ($[0.02,\, 0.44]$)
  & $[0.10,\, 0.33]$ ($[0.02,\, 0.43]$)
  \\
  \hline
  \end{tabular}

  \end{center}

  \renewcommand{\arraystretch}{1.0}
  \caption{
     \label{tab:cTT5:1-dimCLs}
     The constraints on complex-valued $\wilson{T,\,T5}$ --- [upper] 1D-marginalized
     at 68\% (95\%) probability and [lower] 2D-marginalized --- when using
     measurements of $i)$ only $F_H^\mu$, $ii)$ $F_H^\mu$ and other data in
     \reftab{tab:observables} (except $\overline{\cal B}(B_s \to \bar\mu\mu)$), and $iii)$ allowing
     also complex-valud new physics contributions to $\wilson{9,\,10}$. \fred{For 9,10,S,P,T,T5, the absolute value is in $[0.077,0.269] ([0.020, 0.379])$}
  }
\end{table*}

In a scenario with only tensor complex-valued couplings $\wilson{T,\,
  T5}$, the experimental measurement of $F_H^\mu$ constrains the
combination $|\wilson{T}|^2 + |\wilson{T5}|^2$, up to some small
interference of $\wilson{T}$ with vector couplings
$\wilson{7,\,7',\,9,\,9'}$.  The according 68\% (95\%) 1D-marginalized
probability intervals are listed in the first column of
\reftab{tab:cTT5:1-dimCLs}. As can be seen further in the second
column, the constraints become tighter when utilising all observables
in \reftab{tab:observables}, mainly due to the sensitivity of the
branching ratio of $B\to K^{*} \bar\mu\mu$ to tensor couplings (see
also \reffig{fig:FHvsCT}). The latter stronger bounds are driven by
the new lattice results of $B\to K^*$ form factors that yield too high
predictions compared to the measurements~\cite{Horgan:2013pva}. Since
tensor couplings contribute constructively to ${\cal B}(B\to
K^{*}\bar\mu\mu)$ large values are stronger constrained. In this
scenario with vanishing scalar couplings, current measurements of
$A_{\rm FB}^\mu(B^+\to K^+\bar\mu\mu)$ barely provide any constraint;
cf.~\refeq{eq:AFB-dependence}.

We also perform a fit with non-zero complex-valued $\wilson{9,\,10}$
by allowing uniform priors for $\mbox{Re, Im}(\wilson{i}) \in[-7,\,
7]$, ($i = 9,10$) in order to access the stability of the bounds
against interference. Note that $\wilson{7,7'}$ appears in linear
combinations with $\wilson{9,9'}$ such that its interference with
tensor and scalar couplings is captured implicitly by allowing new
physics in $\wilson{9,9'}$. Thus we fix $\wilson{7,7'}$ to the SM
value without loss of generality. In this case, $F_H^\mu$ by itself
still provides bounds on $|\wilson{T,\,T5}|$ that are weakened by a
factor of two since $F_H^\mu$ does not pose constraints on
$\wilson{9,\,10}$ (in the chosen prior range). Once additional
experimental measurements of \reftab{tab:observables} are taken into
account, the potential destructive effects of new physics in
$\wilson{9,\,10}$ become reduced and almost the same constraints on
$\wilson{T,\,T5}$ are recovered, as shown in the third column in
\reftab{tab:cTT5:1-dimCLs}.  In consequence, the $F_H^\mu$
measurement~\cite{Aaij:2014tfa} of LHCb with 3 fb$^{-1}$ shrinks the
previous bounds \cite{Bobeth:2012vn} on $\wilson{T,\,T5}$ by 50\%.

\fred{comment on last row: SM in all cartesian intervals but not in polar!}

%
%
%--------+---------+---------+---------+---------+---------+---------+---------+
\subsection{Scalar couplings \label{sec:scalar}}

\begin{figure*}
  \begin{center}
    \includegraphics[width=.325\textwidth]{plots/pdf/scSP_FH,scSP_Bsmumu,sc910SP_K_KstarBR_Bsmumu_0,2}
    % \includegraphics[width=.32\textwidth]{plots/pdf/scSP_Bsmumu,scSP_FH,sc910SP_K_KstarBR_Bsmumu_4,6}
    \includegraphics[width=.325\textwidth]{plots/pdf/scSP_FH,sc910SP_K_KstarBR_Bsmumu_0,4}
    \includegraphics[width=.325\textwidth]{plots/pdf/scSP_Bsmumu,sc910SP_K_KstarBR_Bsmumu_2,6}
  \end{center}
  \caption{
    The constraints on complex-valued $\wilson{S,\,S',\,P,\,P'}$
    from $i)$ only $F_H^\mu$~(grey dotted), $ii)$ only $\overline{\cal B}(B_s \to \bar\mu\mu)$
    (blue dashed) and $iii)$ their combination with other data in \reftab{tab:observables}
    as well as non-zero $\wilson[]{9,\,9',\,10,\, 10'}$ (red solid)
    at 68\% (darker) and 95\% (lighter) probability. The constraints on
    Re$(\wilson{P}Ê\pm \wilson{P'})$ are identical to Re$(\wilson{S}Ê\pm \wilson{S'})$,
    apart from a small shift of the contours by $(+0.2,\, +0.15)$. The SM is indicated
    by the black diamond.
  }
  \label{fig:scalar}
\end{figure*}

\begin{table*}
  \renewcommand{\arraystretch}{1.3}
  \begin{center}
  \begin{tabular}{cccc}
  \hline
  data set
  & only $F_H^\mu$
  & only $\overline{\cal B}(B_s \to \bar\mu\mu)$
  & all
  \\
  set of couplings
  & $\wilson{S,\,S',\,P,\,P'}$
  & $\wilson{S,\,S',\,P,\,P'}$
  & $\wilson{S,\,S',\,P,\,P',\,9,\,9',\,10,\,10'}$
  \\
  \hline
  $\mbox{Re}(\wilson{S} - \wilson{S'})$
  & $\quad$ $-$ $\quad$
  & $\quad$ $[-0.10,\, 0.08]$ ($[-0.14,\, 0.13]$) $\quad$
  & $\quad$ $[-0.08,\, 0.07]$ ($[-0.13,\, 0.13]$) $\quad$
  \\
  $\mbox{Im}(\wilson{S} - \wilson{S'})$
  & $\quad$ $-$ $\quad$
  & $\quad$ $[-0.07,\, 0.07]$ ($[-0.11,\, 0.12]$) $\quad$
  & $\quad$ $[-0.07,\, 0.07]$ ($[-0.12,\, 0.11]$) $\quad$
  \\[0.1cm]
  $\mbox{Re}(\wilson{S} + \wilson{S'})$
  & $[-0.36,\, 0.39]$ ($[-0.69,\, 0.68]$)
  & $-$
  & $[-0.32,\, 0.32]$ ($[-0.59,\, 0.62]$)
  \\
  $\mbox{Im}(\wilson{S} + \wilson{S'})$
  & $[-0.37,\, 0.35]$ ($[-0.68,\, 0.66]$)
  & $-$
  & $[-0.25,\, 0.41]$ ($[-0.57,\, 0.64]$)
  \\[0.1cm]
  $\mbox{Re}(\wilson{P} - \wilson{P'})$
  & $-$
  & $[ 0.05,\, 0.20]$ ($[ 0.01,\, 0.26]$)
  & $[ 0.00,\, 0.16]$ ($[-0.07,\, 0.22]$)
  \\
  $\mbox{Im}(\wilson{P} - \wilson{P'})$
  & $-$
  & $[-0.07,\, 0.08]$ ($[-0.12,\, 0.12]$)
  & $[-0.07,\, 0.09]$ ($[-0.14,\, 0.16]$)
  \\[0.1cm]
  $\mbox{Re}(\wilson{P} + \wilson{P'})$
  & $[-0.24,\, 0.51]$ ($[-0.51,\, 0.82]$)
  & $-$
  & $[-0.12,\, 0.52]$ ($[-0.42,\, 0.78]$)
  \\
  $\mbox{Im}(\wilson{P} + \wilson{P'})$
  & $[-0.36,\, 0.37]$ ($[-0.67,\, 0.67]$)
  & $-$
  & $[-0.37,\, 0.29]$ ($[-0.68,\, 0.57]$)
  \\
  \hline
  \end{tabular}
  \end{center}

  \renewcommand{\arraystretch}{1.0}
  \caption{
     \label{tab:cSP:1-dimCLs}
     The 1D-marginalized constraints on complex-valued $\wilson{S,\,S',\,P,\,P'}$
     at 68\% (95\%) probability from measurements of $i)$ only $F_H^\mu$,
     $ii)$ only $\overline{\cal B}(B_s \to \bar\mu\mu)$ and $iii)$ all the data in
     \reftab{tab:observables} and additional new physics contributions to
     $\wilson{9,\, 9',\, 10,\,10'}$.
  }
\end{table*}

Scalar couplings $\wilson{S,\, S',\, P,\, P'}$ enter $F_H^\ell$
without kinematic suppression --- see \refeq{eq:FH-dependence} --- in
the complementary manner $(\wilson{i} + \wilson{i'})$ to the branching
ratio $\overline{\cal B}(B_s\to \bar\ell\ell)$, which constrains only
$(\wilson{i} - \wilson{i'})$, where $i = S,\, P$.  Most notably, we
find that the current measurement of $F_H^\mu$ does constrain the
orthogonal $(\wilson{i} + \wilson{i'})$, such that the combination of
$F_H^\mu$ and $\overline{\cal B}(B_s \to \bar\mu\mu)$ allows to bound
for the first time the real and imaginary parts of all four couplings.

The according four-parameter fit for complex-valued scalar couplings
to the data of $\overline{\cal B}(B_s \to \bar\mu\mu)$ and $F_H^\mu$
results in the (2D-marginalized) regions in the $\mbox{Re}(\wilson{i}
\pm \wilson{i'})$ ($i = S, P$) planes shown in
\reffig{fig:scalar}. The corresponding plots for $\mbox{Im}(\wilson{i}
\pm \wilson{i'})$ are very similar to those shown and thus
omitted. These bounds do not change when including all other data in
\reftab{tab:observables}, since the $A_{\rm FB}^\mu(B\to K\bar\mu\mu)$
requires interference of scalar with tensor couplings and other
observables are not very sensitive to scalar
couplings. Quantitatively, the constraint from $\overline{\cal B} (B_s
\to \bar\mu\mu)$ on $(\wilson{i} - \wilson{i'})$ is about a factor
four to five stronger than the one of $F_H^\mu$ on $(\wilson{i} +
\wilson{i'})$.

Interference terms of $\wilson{P,\,P'}$ with vector couplings might
weaken these bounds. Most important for $\overline{\cal B}(B_s \to
\bar\mu\mu)$ are $(\wilson{10} - \wilson{10'})$ --- compare
\refeq{eq:Br-Bsmumu:time0} --- and for $F_H^\ell$ similarly
$(\wilson{10} + \wilson{10'})$ \cite{Bobeth:2007dw}, which are
suppressed by factors $m_\mu/M_{B_s}$ and $m_\mu/\sqrt{q^2}$,
respectively \fred{hard to read}. However, these terms become
important for small $\wilson{P,\,P'}$ due to the large SM contribution
of $\wilson[SM]{10} \simeq -4.2$. We compile bounds on complex-valued
scalar couplings in \reftab{tab:cSP:1-dimCLs} for only $F_H^\mu$, only
$\overline{\cal B}(B_s \to \bar\mu\mu)$, and their combination with
all the other observables in \reftab{tab:observables}. In the last
case, we allow new physics in vector couplings $\wilson{9,\ 9',\,
  10,\, 10'}$ assuming uniform priors for $\mbox{Re, Im}(\wilson{i})
\in[-7,\, 7]$.  Neither $F_H^\ell$ nor $\overline{\cal B}(B_s \to
\bar\ell\ell)$ alone can bound all four complex-valued scalar
couplings, however their combination is capable to do so and moreover,
the bounds are stable against destructive interference with vector
couplings. In the special case of real-valued couplings,
$\overline{\cal B}(B_s \to \bar\ell\ell)$ would lead to
rings~\cite{Alonso:2014csa} instead of circles. In the most general
scenario, we find the following bounds with 68\% (95\%) probability:
\begin{equation}
  \begin{aligned}
    \left| \wilson{S} \right| & \in [0.06,\, 0.25]\; ([0.01,\, 0.36]),\\
    \left| \wilson{S'} \right| & \in [0.09,\, 0.46]\; ([0.02,\, 0.73]),\\
    \left| \wilson{P} \right| & \in [0.05,\, 0.21]\; ([0.01,\, 0.32]),\\
    \left| \wilson{P'} \right| & \in [0.11,\, 0.42]\; ([0.02,\, 0.63]).
  \end{aligned}
\end{equation}

Our results improve and extend previous bounds in the literature to
the most general case of complex-valued couplings. For example they
are a factor two to five more stringent than \cite{Becirevic:2012fy}
and comparable to \cite{Altmannshofer:2012az} once restricting to the
much simpler scenarios considered there.

\begin{figure*}
  \begin{center}
      \includegraphics[width=.35\textwidth]{plots/pdf/smEFTS_K_Bsmumu,smEFTS10_K_KstarBR_Bsmumu_0,2}
      \includegraphics[width=.35\textwidth]{plots/pdf/smEFTS_K_Bsmumu,smEFTS10_K_KstarBR_Bsmumu_1,3}
    \end{center}
    \caption{
    Constraints (2D-marginalized) on complex-valued scalar couplings
    $\wilson{S,\,S'}$ in the scenario SM-EFT from all data in
    \reftab{tab:observables} (red) at 68\% (darker) and 95\%
    (lighter) probability when marginalising over non-zero $\mbox{Re, Im}
    (\wilson[NP]{10,\,10'}) \in [-7,\,7]$.
    For comparison the according constraints from only $\overline{\cal B}(B_s\to \bar\mu\mu)$
    and $F_H^\mu$ \todo{What else?}
    (dashed blue contours) at 68\% (inner) and 95\% (outer) probability.
    The SM is indicated by the black diamond.
  }
  \label{fig:scalar:SM-EFT}
\end{figure*}

%
%
%--------+---------+---------+---------+---------+---------+---------+---------+
\subsection{SM-EFT-constrained scalar couplings \label{sec:SM-EFT}}

In the following we consider a model-independent scenario as predicted by the SM
augmented by dimension six operators that respect the SM gauge group and are
composed out of SM fields only. After the discovery of a scalar resonance at the LHC
that complies currently with all requirements of the Higgs particle in the SM
and the steadily rising lower bounds on new particle masses towards higher energies
from ATLAS and CMS such a scenario becomes more and more viable \fred{hard to read}.

A non-redundant set of dimension-six operators of this effective
theory (SM-EFT) was given in \cite{Grzadkowski:2010es} and the
matching to the effective theory of $\Delta B = 1$ decays
\refeq{eq:Heff} at the scale $\mu \sim m_W$ of the order of the
$W$-boson mass in~\cite{D'Ambrosio:2002ex} for vector couplings
$\wilson{7,\,9,\,10}$. The matching for tensor \refeq{eq:tensor:ops}
and scalar \refeq{eq:scalar:ops} operators \cite{Alonso:2014csa} shows
that SM gauge groups in conjunction with the linear representation
\todo{Ask Martin, Oscar how to correctly phrase this} impose the
relations
\begin{align}
  \label{eq:mEFT-constraints}
  \wilson{P}  & = -\wilson{S}\,, &
  \wilson{P'} & = \wilson{S'}\,, &
  \wilson{T}  & = \wilson{T5} = 0
\end{align}
on scalar and pseudo-scalar couplings and require tensor couplings to be
suppressed to the level of dimension eight operators. In consequence only two
scalar couplings $\wilson{S,\, S'}$ arise that scale as $\sim v^2/\Lambda_{\rm NP} \ll 1$
where $v^2$ denotes the scale of electroweak symmetry breaking and $\Lambda_{\rm NP}$
a high energy scale at which new physics generates these operators.

The couplings $\wilson{S,\, S'}$ can be bound quite reliably from (omitting
for the sake of the argument terms of order $m_\ell^2/M_{B_s}^2$ and $m_s/m_b$)
\cite{Alonso:2014csa}
\begin{equation}
\begin{aligned}
  \overline{\cal B}(B_s\to \bar\ell\ell) & \propto
      \left|\wilson{S} + \wilson{S'}\right|^2
    + \left|\wilson{S} - \wilson{S'}\right|^2
\\[0.2cm]
  - & \frac{4 m_\ell\, m_b}{M_{B_s}^2} \mbox{Re}
  \left[ (\wilson{S} + \wilson{S'})(\wilson{10} - \wilson{10'})^* \right]
\end{aligned}
\end{equation}
and almost similarly (omitting terms of order $m_\ell^2/q^2$
and $m_s/m_b$)
\begin{equation}
\begin{aligned}
  F_H^\ell & \propto
      \left|\wilson{S} + \wilson{S'}\right|^2
    + \left|\wilson{S} - \wilson{S'}\right|^2
\\[0.2cm]
  & - \frac{4 m_\ell\, m_b}{q^2} \mbox{Re}
  \left[ (\wilson{S} - \wilson{S'})(\wilson{10} + \wilson{10'})^* \right] \,.
\end{aligned}
\end{equation}

The SM gauge group does not imply any relation between $\wilson{10}$
and $\wilson{10'}$, so they are in general additional independent
parameters. Here we find that destructive interference with
contributions involving $\wilson{10,\, 10'}$ does not significantly
alter the bounds on $\wilson{S,\, S'}$.  Since we consider here
complex-valued couplings the allowed regions are circles rather than
rings as for the case of real-valued
couplings~\cite{Alonso:2014csa}. The result of the simultaneous fit of
complex-valued $\wilson{S,\,S',\,10,\,10'}$ is shown in
\reffig{fig:scalar:SM-EFT}. \fred{Circle smaller, prob. moves to
  center compared to Alonso}

It must be noted that relations \refeq{eq:mEFT-constraints} are a consequence
of embedding the Higgs in a weak doublet along with the Goldstone bosons.
For example choosing a non-linear representation of the scalar sector allows
for additional dimension six operators in the SM-EFT, such that $\wilson{S,S',P,P'}$
are all independent and tensor operators have non-vanishing couplings already
at dimension six \cite{Cata:2015lta}.

%
%
%--------+---------+---------+---------+---------+---------+---------+---------+
\subsection{Tensor, scalar and vector couplings \label{sec:tensor:scalar}}

The most general fit of complex-valued tensor and scalar couplings
$\wilson{S,S',P,P',T,T5}$ in combination with complex-valued vector
couplings $\wilson{9,\,9',\,10,\, 10'}$ --- i.e. the combination of
\refsec{sec:tensor} and \refsec{sec:scalar} with same prior ranges ---
yields bounds very similar to those in tables \ref{tab:cTT5:1-dimCLs}
and \ref{tab:cSP:1-dimCLs} . The changes are only small and in fact
the bounds tend to be even more stringent because tensor and scalar
couplings can contribute to $F_H^{\mu}$ only constructively --- see
\refeq{eq:FH-dependence}.  As before, branching-ratio measurements of
$B\to K^{*}\bar\mu\mu$ help to strengthen the constraints on
$\wilson{T,\, T5}$. This demonstrates that even in the case of
complex-valued couplings the data are able to constrain all real and
imaginary parts.

%
%
%--------+---------+---------+---------+---------+---------+---------+---------+
\subsection{Angular observables in $B\to K^* \bar\ell\ell$}

\begin{table*}
  \renewcommand{\arraystretch}{1.5}
  \begin{center}
  \begin{tabular}{cccccc}
  \hline
    observable
  & $q^2$-bin [GeV$^2$]
  & SM
  & $\quad T(5),\,9,\,10\quad$
  & $\quad S^{(')},\,P^{(')},\,9^{(')},\,10^{(')} \quad$
  & $\quad S^{(')},P^{(')},\,T(5),\,9^{(')},10^{(')}$
  \\
  \hline
  \multirow{2}{*}{$\tau_{B^0} \times J_{6c}$}
  & $[1.1,\, 6]$
  & $\simeq 0$
  & $(0.7_{-2.1}^{+1.7}) \cdot 10^{-9\;}$
  & $(0.0_{-2.1}^{+2.2}) \cdot 10^{-\todo{10}}$
  & $(0.2_{-1.1}^{+1.2}) \cdot 10^{-9}$
  \\
  & $[15,\, 19]$
  & $\simeq 0$
  & $(2.1_{-6.2}^{+5.1}) \cdot 10^{-10}$
  & $(0.7_{-6.6}^{+5.4}) \cdot 10^{-11}$
  & $(-0.2_{-2.9}^{+3.5}) \cdot 10^{-11}$
  \\
  \hline
  \multirow{2}{*}{$\tau_{B^0} \times (J_{1c} + J_{2c})$}
  & $[1.1,\, 6]$
  & $(3.54_{-0.66}^{+0.65}) \cdot 10^{-9\;}$
  & $(5.0_{-1.9}^{+1.7}) \cdot 10^{-9\;}$
  & $(2.9_{-0.5}^{+0.4}) \cdot 10^{-9\;}$
  & $(2.8_{-1.0}^{+1.6}) \cdot 10^{-9}$
  \\
  & $[15,\, 19]$
  & $(1.72_{-0.16}^{+0.16}) \cdot 10^{-10}$
  & $(4.7_{-2.6}^{+3.3}) \cdot 10^{-9\;}$
  & $(1.9_{-0.7}^{+0.3}) \cdot 10^{-10}$
  & $(3.1_{-2.3}^{+2.5}) \cdot 10^{-9}$
  \\
  \hline
  \multirow{2}{*}{$\tau_{B^0} \times (J_{1s} - 3 J_{2s})$}
  & $[1.1,\, 6]$
  & $(2.98_{-0.54}^{+0.63}) \cdot 10^{-10}$
  & $(1.9_{-1.1}^{+1.4}) \cdot 10^{-8\;}$
  & $(4.9_{-1.4}^{+1.6}) \cdot 10^{-10}$
  & $(1.6_{-1.2}^{+0.9}) \cdot 10^{-8}$
  \\
  & $[15,\, 19]$
  & $(1.12_{-0.10}^{+0.10}) \cdot 10^{-10}$
  & $(7.1_{-3.8}^{+5.6}) \cdot 10^{-9\;}$
  & $(1.4_{-0.4}^{+0.3}) \cdot 10^{-10}$
  & $(5.3_{-3.8}^{+3.8}) \cdot 10^{-9}$
  \\
  \hline
  \multicolumn{2}{c}{$A_{\Delta \Gamma}(B_s \to \bar\mu\mu)$}
  & 1
  & SM
  & $[-1,\, 1]$
  & $[-1,\, 1]$
  \\
  \hline
  \end{tabular}
  \end{center}
  \renewcommand{\arraystretch}{1.0}
  \caption{
     \label{tab:postdictions}
     The 68\% probability intervals of not yet measured angular observables in
     $B^0\to K^{*0}\bar\mu\mu$ integrated over $q^2$ bins at low- and high-$q^2$
     for several new physics scenarios respecting all the considered experimental
     constraints. The SM prior predictives are given too, where ``$\simeq 0$''
     indicates zero in the considered
     approximation, see text for more explanations.
  }
\end{table*}

Now we present the yet allowed ranges due to non-standard tensor and scalar
couplings of the so far unmeasured angular observables $J_{6c}$, $(J_{1s} - 3\,
J_{2s})$ and $(J_{1c} + J_{2c})$ in $B\to K^*(\to K\pi)\bar\ell\ell$ in order to
explore the potential of their future measurements in fits of the kind pursued
in this work.  The 68\% probability intervals are collected in
\reftab{tab:postdictions} for the various model-independent scenarios, allowing
for interference with non-zero vector couplings, in \refsec{sec:tensor},
\refsec{sec:scalar} and \refsec{sec:tensor:scalar} for $q^2$-integrated
quantities in the low- and high-$q^2$ bins $q^2\in [1.1,\, 6]$ GeV$^2$ and
$q^2\in [15,\, 19]$ GeV$^2$, respectively. Their posterior distributions have
been obtained from the posterior samples as explained in~\ref{app:mc}.

For comparison, the SM estimates are also given in \reftab{tab:postdictions}
multiplied by the $B^0$-meson life time $\tau_{B^0} = 1.519$~ps.
At low-$q^2$ they are of order ${\cal O}(10^{-9})$ for $(J_{1s} - 3\, J_{2s})$ and
$(J_{1c} + J_{2c})$, getting suppressed down to ${\cal O}(10^{-10})$ at high-$q^2$
due to the $m_\ell/\sqrt{q^2}$ dependence. In order to compare it to the typical
magnitude of vector couplings, note that for example the branching ratio of
$B\to K^*\bar\ell\ell$
\begin{align}
  \frac{{\rm d}\BR}{{\rm d}q^2} & = \frac{\tau_{B^0}}{3} \left[
    2\, (3 J_{1s} - J_{2s}) + (3 J_{1c} -  J_{2c})
  \right]
\end{align}
is approximately equal to $2 \cdot 10^{-7}$ in the SM for both the
$q^2 \in [1,\, 6]$ and $[15,\, 19]$ GeV$^2$ bins.

The angular observable $J_{6c}$ is strictly zero in the absence of
tensor and scalar couplings. Non-zero contributions can be generated
in the SM by QED corrections or potentially from higher-dimensional
($d \geq 8$) $|\Delta B| = |\Delta S| = 1$ operators, leading to
parametrical suppression of $\alpha_e/(4 \pi)$ or $(m_b/m_W)^2$,
respectively. Such a suppression must be compared to potential
suppression present for tensor and scalar contributions in particular
NP models in order to decide their relevance. Our model-independent
fits are still in a realm where such considerations are insignificant
since current experimental measurements, in combination with theory
uncertainties, do not impose yet too stringent constraints on tensor
and scalar couplings.

Beyond the SM, $J_{6c}$ can become of order ${\cal O}(10^{-9})$ in
scenarios involving tensor couplings only and about ${\cal
  O}(10^{-10})$ in the presence of scalar couplings only. Both effects
are due to interference with vector couplings. Due to the strong
suppression of scalar couplings, the direct interference of scalar and
tensor contributions the allowed region in the most general scenario
\todo{schematic formula from notes}

Concerning $(J_{1c} + J_{2c})$ and $(J_{1s} - 3\, J_{2s})$, again substantial
deviations from the SM prediction are possible only in the presence of non-zero
tensor couplings. In this case enhancement by two orders of magnitude are
possible up to the ${\cal O}(10^{-8})$ at high-$q^2$ and also at low-$q^2$ in
the case of $(J_{1s} - 3\, J_{2s})$, always being consistent with all included
experimental constraints in \reftab{tab:observables}. In view of current
experimental precision of $20\%$ of the branching ratio measurements at LHCb
\cite{Aaij:2013iag} with only 1 fb$^{-1}$, corresponding to the ${\cal
  O}(10^{-8})$, one can indeed hope for some sensitivity to such large effects
in $(J_{1c} + J_{2c})$ and $(J_{1s} - 3\, J_{2s})$ for the not yet published 3
fb$^{-1}$ data set at least with the method of moments~\cite{beaujean:2015mom}.

The CP asymmetry $A_{\Delta\Gamma}(B_s\to \bar\mu\mu)$ induced by
non-vanishing width of the $B_s$ meson (cf. \refapp{app:BKstarellell})
can still take any value in the range $[-1,\, 1]$ in scenarios with
nonzero scalar couplings.

%
%
%
%--------+---------+---------+---------+---------+---------+---------+---------+
\section{Conclusions}

We have derived to date most stringent constraints on tensor and scalar
couplings that mediate $b\to s\bar\mu\mu$ utilising the latest measurements
of angular observables $F_H^\mu$ and the lepton forward-backward asymmetry
$A_{\rm FB}^\mu$ in $B^+\to K^+ \bar\mu\mu$ from LHCb \cite{Aaij:2014tfa},
supplemented by measurements of the branching ratios of $B_s\to \bar\mu\mu$ and
$B\to K^{(*)} \bar\mu\mu$. Both observables, $F_H^\mu$ and $A_{\rm FB}^\mu$,
belong to a class in which vector and dipole couplings that arise also in
the standard model (SM) are suppressed (mostly kinematically by $m_\ell /
\sqrt{q^2}$) w.r.t tensor and scalar couplings. We provide predictions for
analogous but not yet measured angular observables $J_{6c}$, $(J_{1c} + J_{2c})$
and $(J_{1s} - 3 J_{2s})$ in $B\to K^* (\to K\pi) \bar\mu\mu$.

We find that the experimental measurement of $F_H^\mu$, especially at high-$q^2$,
\begin{enumerate}
\item imposes by itself constraints on tensor couplings $|\wilson{T,\,T5}|
  < 0.43\, (0.57)$ at 68\% (95\%) probability, superseding previous bounds.
  In combination with current data from $B\to K^* \bar\mu\mu$ and lattice
  $B\to K^*$ form factors they become tighter $|\wilson{T,\,T5}| < 0.33\,
  (0.43)$, even in the presence of non-standard contributions in vector
  couplings.
\item allows for the first time to bound all four scalar couplings $\wilson{i}$
  with $i = S,\,S',\,P,\,P'$ due to it's complementarity to $\overline{\cal B}(B_s
  \to \bar\mu\mu)$. Even when taking into account destructive interference with
  vector couplings, $|\wilson{i} + \wilson{i'}| < 0.3\, (0.6)$ and
  $|\wilson{i} - \wilson{i'}| < 0.1\, (0.2)$ for $i = S,P$ at 68\% (95\%)
  probability. Currently, the bounds from $F_H^\mu$ are not as strong as those
  from $\overline{\cal B}(B_s \to \bar\mu\mu)$ by about a factor four. In the
  future, measurements at LHCb and Belle II will further tighten the bounds.

  Moreover, measurements of $F_H^e$ ($\ell = e$)
  will provide constraints on scalar couplings in the electron channel in
  the lack of measurements of the branching ratio of $B_s \to \bar{e}e$.
\end{enumerate}
Our updated bounds on complex-valued tensor and scalar couplings are summarised
in \reftab{tab:cTT5:1-dimCLs} and \reftab{tab:cSP:1-dimCLs}, accounting also for
interference effects with vector couplings. These bounds hold even in the most
general scenario of complex-valued tensor, scalar and vector couplings, showing
also that the data is able to bound the real and imaginary parts of all Wilson
coefficients simultaneously.

As a special case, we considered also the scenario arising from the SM augmented
by dimension 6 operators generalising existing studies to the case of complex-valued
couplings. In this scenario tensor couplings are absent and additional
relations between scalar couplings are enforced by the linear realisation of the
$SU(2)_L \otimes U(1)_Y$ electroweak symmetry group.

Our study of yet not measured angular observables $J_{6c}$, $(J_{1c} + J_{2c})$
and $(J_{1s} - 3 J_{2s})$ in $B\to K^* (\to K\pi) \bar\mu\mu$ (see
\reftab{tab:postdictions}) shows that despite the current bounds on tensor couplings,
enhancements of two orders of magnitude over the SM predictions are allowed for
$(J_{1c} + J_{2c})$ and $(J_{1s} - 3 J_{2s})$. On the other hand, current bounds
on scalar couplings from $B_s\to \bar\mu\mu$ and $F_H^\mu$ are already quite
restrictive allowing only for small deviations from SM predictions in
$(J_{1c} + J_{2c})$ and $(J_{1s} - 3 J_{2s})$. Notably, the CP asymmetry
$A_{\Delta\Gamma}(B_s\to \bar\mu\mu)$ induced by non-vanishing width of the
$B_s$ meson can still take any value in the range $[-1,\, 1]$.

%
%
%
%--------+---------+---------+---------+---------+---------+---------+---------+
\begin{acknowledgements}
  We are grateful to Danny van Dyk for helpful discussions and his support on
  EOS \cite{EOS}. Concerning form factor results from lattice, we thank Chris
  Bouchard and Matthew Wingate for communications on their results of $B\to K$
  \cite{Bouchard:2013eph} and $B\to K^*$ \cite{Horgan:2015vla} form factors,
  respectively. We thank also David Straub for his support with LCSR results of
  $B\to K^*$ form factors \cite{Straub:2015ica}.
  C.B. was supported by the ERC Advanced Grant project ``FLAVOUR''
  (267104).
\end{acknowledgements}

%--------+---------+---------+---------+---------+---------+---------+---------+
%
%  Appendix
%
%--------+---------+---------+---------+---------+---------+---------+---------+

\appendix

%
%
%
%--------+---------+---------+---------+---------+---------+---------+---------+
\section{
  Angular observables in $B\to K^* \bar\ell\ell$
  \label{app:BKstarellell}
}

Here we focus on those angular observables $J_i$ in $B\to K^* \bar\ell\ell$ in
which tensor and scalar contributions are kinematically enhanced by a factor
$\sqrt{q^2}/m_\ell$ compared to the vector contributions of the SM or the
respective interference terms. The results and the notation follow
\cite{Bobeth:2012vn}.

For convenience, we split the time-like transversity amplitude $A_t$ into two
parts as
\begin{align}
  A_t &
  = \tilde{A}_t - \frac{1}{2} \frac{\sqrt{q^2}}{m_\ell} A_P \, .
\end{align}
Both terms depend on the scalar $B\to K^*$ form factor $A_0(q^2)$, a
normalization factor $N$, and the K{\"a}ll{\'e}n-function $\lambda$ (see
\cite{Bobeth:2012vn}) but have a different dependence on the Wilson coefficients
$i = 10, 10', P, P'$:
\begin{equation}
\begin{aligned}
  \tilde{A}_t & = 2 N \frac{\sqrt{\lambda}}{\sqrt{q^2}}
    (\wilson[]{10} - \wilson[]{10'}) A_0\,,
\\
  A_P & = - 2 N\sqrt{\lambda} \frac{(\wilson[]{P} - \wilson[]{P'})}{(m_b+m_s)} A_0\,.
\end{aligned}
\end{equation}
The lepton-flavor index $\ell$ of Wilson coefficients is omitted for brevity
throughout.

In full generality, the angular observables in $B\to K^* \bar\ell\ell$ depend on
seven transversity amplitudes with vector and dipole contributions,
$A_{0,\,\perp,\,\parallel}^{L,R}$ and $\tilde{A}_t$, one scalar and one
pseudoscalar amplitude, $A_{S,\,P} \propto (\wilson[]{S,P} - \wilson[]{S',P'})$,
and six tensor amplitudes, $A_{\parallel\perp,\,t\perp,\,0\parallel} \propto
\wilson[]{T}$ and $A_{t0,\,t\parallel,\,0\perp} \propto \wilson[]{T5}$.  The
interesting combinations are
\begin{align}
  \frac{4}{3} J_{6c}  =
  4\, \beta_\ell\, \mbox{Re} \bigg[ &2\, (A_{t0}^{} A_S^* - A_{\parallel\perp}^{} A_P^*) \nonumber\\
  &+ \frac{m_\ell}{\sqrt{q^2}} \big[(A_0^L + A_0^R) A_S^*
           + 4\, A_{\parallel\perp}^{} \tilde{A}_t^* \big] \bigg],
\end{align}
\begin{align}
  \label{eq:j1c-j2c}
  \frac{4}{3}\(J_{1c} + J_{2c}\) = \;
  &\Bigg| \frac{2\, m_\ell}{\sqrt{q^2}} (A_0^L + A_0^R) + 4\, A_{t0} \Bigg|^2    + 16\, \beta_\ell^2\, |A_{\parallel\perp}|^2 \nonumber\\
  + &\Bigg|\frac{2\, m_\ell}{\sqrt{q^2}}\, \tilde{A}_t - A_P\Bigg|^2 +
  \beta_\ell^2\, |A_S|^2,
 \end{align}
 \begin{align}
 \frac{4}{3}\(J_{1s} - 3\, J_{2s}\) = \;
    &\Bigg| \frac{\sqrt{2}\, m_\ell}{\sqrt{q^2}} (A_\perp^L + A_\perp^R) + 4\, A_{t\perp} \Bigg|^2 \nonumber\\
  + &\Bigg| \frac{\sqrt{2}\, m_\ell}{\sqrt{q^2}} (A_\parallel^L + A_\parallel^R) + 4\, A_{t\parallel} \Bigg|^2.
\end{align}
The function $\beta_\ell^2(q^2) \equiv 1 - 4 m_\ell^2/q^2$ tends to 1 for $m_\ell \ll
\sqrt{q^2}$.  This condition is well fulfilled for $\ell = e$ and $q^2 \gtrsim
1\, \mbox{GeV}^2$, provided that tensor and scalar Wilson coefficients do not
receive additional suppression factors. For $\ell = \mu$, the value of $q^2$
should not be too low, whereas in the case $\ell = \tau$, these observables are
not anymore dominated by tensor and scalar contributions alone, and the full
lepton-mass dependence has to be taken into account.  Finally, we note that the
second part of $(J_{1c} + J_{2c})$ in \refeq{eq:j1c-j2c},
\begin{align}
  &\Bigg|\frac{2\, m_\ell}{\sqrt{q^2}}\, \tilde{A}_t - A_P\Bigg|^2
   + \beta_\ell^2\, |A_S|^2   =  4 N^2 \lambda A_0^2 \Bigg\{
    \beta_\ell^2 \Bigg|\frac{\wilson[]{S} - \wilson[']{S}}{m_b + m_s}\Bigg|^2\nonumber\\
  &\hskip1cm + \Bigg|\frac{\wilson[]{P} - \wilson[']{P}}{m_b + m_s}
     + \frac{2 m_\ell}{q^2} \big(\wilson[]{10} - \wilson[']{10}\big) \Bigg|^2
    \Bigg\}\,,
\end{align}
 resembles very much the branching ratio of the rare
decay $B_s \to \bar\ell\ell$ in the limit $q^2 \to M_{B_s}^2$
\begin{align}
\label{eq:Br-Bsmumu:time0}
  {\cal B}(B_s \to \bar\ell\ell) &=
    \frac{G_F^2 \alpha_e^2 |V_{tb}^{} V_{ts}^*|^2}{64\, \pi^3}
    M_{B_s}^5 f_{B_s}^2 \tau_{B_s} \beta_\ell(q^2 = M_{B_s}^2)
\nonumber\\
    & \times \Bigg\{
    \beta_\ell^2(q^2 = M_{B_s}^2) \Bigg|\frac{\wilson[]{S} - \wilson[']{S}}{m_b + m_s}\Bigg|^2 \nonumber\\
    &+ \Bigg|\frac{\wilson[]{P} - \wilson[']{P}}{m_b + m_s}
     + \frac{2 m_\ell}{M_{B_s}^2} \big(\wilson[]{10} - \wilson[']{10}\big) \Bigg|^2
    \Bigg\}\,.
  \end{align}
  Hence there is some similarity between $(J_{1c} + J_{2c})$ in $B \to K^*
  \bar\ell\ell$ and ${\cal B}(B_s \to \bar\ell\ell)$ in their dependence on the
  couplings but the former has additional dependence on tensor and vector
  couplings through the other transversity amplitudes.  In $B \to K^*
  \bar\ell\ell$, the helicity suppression factor $4\,m_\ell^2/q^2$ of vector
  couplings is weaker than the corresponding factor $4\,m_\ell^2/M_{B_s}^2$ in
  $B_s \to \bar\ell\ell$.

  Concerning $B_s\to \bar\ell\ell$, the expression \refeq{eq:Br-Bsmumu:time0}
  corresponds to the unintegrated branching ratio at time $t=0$. Due to the
  non-vanishing decay width $\Delta\Gamma_s$, experiments measure the average
  time-integrated branching ratio---denoted by $\overline{\cal B}$---and the two
  are related as \cite{DeBruyn:2012wk}
\begin{align}
  \label{eq:Br-Bsmumu}
  \overline{\cal B} (B_s \to \bar\ell\ell) &
  = \frac{1 + y_s {\cal A}_{\Delta\Gamma}}{1 - y_s^2} {\cal B} (B_s \to \bar\ell\ell)\,.
\end{align}
Here $y_s \equiv \Delta\Gamma_s/(2\Gamma_s)$ with the numerical value given in
\cite{Beaujean:2013soa}. ${\cal A}_{\Delta\Gamma}$ is the CP asymmetry
due to nonvanishing width difference, which is ${\cal A}_{\Delta\Gamma} = 1$
in the SM, but in general can be ${\cal A}_{\Delta\Gamma} \in [-1,\, 1]$.
Since ${\cal A}_{\Delta\Gamma}$ can depart from it's SM value in scenarios of
new physics considered in this work, we take this effect into account
in our numerical analysis, although it is suppressed by small $y_s$. The latest
SM prediction $\overline{\cal B} (B_s \to \bar\mu\mu) = (3.65 \pm 0.23) \cdot
10^{-9}$~\cite{Bobeth:2013uxa} includes NLO electroweak \cite{Bobeth:2013tba}
and NNLO QCD corrections \cite{Hermann:2013kca}.

Finally we discuss the possibility of suitable normalizations of $J_{6c}$,
$(J_{1s} - 3\, J_{2s})$ and $(J_{1c} + J_{2c})$ at high $q^2$ that would provide
optimized observables. For this purpose we use form-factor relations at leading
order in $1/m_b$ and neglect terms suppressed by $m_\ell/\sqrt{q^2}$. With the
notation and expressions derived in \cite{Bobeth:2012vn},
\begin{equation}
\begin{aligned}
  J_{1s} - 3\, J_{2s} & = \frac{3}{2}\, \rho_1^T (f_\perp^2 + f_\parallel^2) \,,
\\[0.1cm]
  J_{1c} + J_{2c} & = 3\, \rho_1^T f_0^2
     + \frac{3 N^2 \lambda}{(m_b+m_s)^2} A_0^2
\\
   & \quad \times \left(|\wilson{S}-\wilson{S'}|^2 + |\wilson{P}-\wilson{P'}|^2\right)\,.
%\\[0.1cm]
%  \frac{4}{3} J_{6c} & = \todo{???}
\end{aligned}
\end{equation}
Here $f_{\perp, \parallel, 0}$ and $A_0$ denote $B\to K^*$ form factors, whereas
the $\rho_1^\pm$ depends on vector couplings and $\rho_1^T \propto
(|\wilson{T}|^2 + |\wilson{T5}|^2)$.

Concerning $(J_{1s} - 3\, J_{2s})$, there are no appropriate normalizations,
unless the chirality-flipped $\wilson[NP]{7',\,9',\,10'} = 0$ because then
$\rho_1^+ = \rho_1^-$ holds. There are three potential normalizations
\begin{equation}
\begin{aligned}
  \frac{4}{3} & J_{1s} =
    (3 \rho_1^+ + \rho_1^T) f_\perp^2 +  (3 \rho_1^- + \rho_1^T) f_\parallel^2 \,,
\\
  \frac{8}{3} & J_{2s} =
    (\rho_1^+ - \rho_1^T) f_\perp^2 +
    (\rho_1^- - \rho_1^T) f_\parallel^2 \,,
\\[0.1cm]
  & J_{1s} + 2 J_{2s} =
    3 (\rho_1^+ f_\perp^2 + \rho_1^- f_\parallel^2) \,
\end{aligned}
\end{equation}
where the last one depends only on vector couplings; i.e., is free of tensor and
scalar ones. In $J_{1s}$ and $J_{2s}$, tensor couplings contribute either
cumulatively or destructively to vector couplings in $\rho_1^\pm$.

A similar situation arises for $(J_{1c} + J_{2c})$, where no appropriate
normalization exists, unless scalar couplings vanish. In this special case, both
$J_{1c}$ and $J_{2c} = -3/2 (\rho_1^- - \rho_1^T) f_0^2$ depend only on $f_0$
and can be used as normalization. In the general case, $J_{2c}$ still depends
only on tensor couplings and was used at low $q^2$ \cite{Matias:2012xw}. Finally
we note that
\begin{equation}
\begin{aligned}
  J_{1c} - J_{2c} & = 3\,\rho_1^- f_0^2 + \frac{3 N^2 \lambda}{(m_b+m_s)^2} A_0^2
\\
   & \quad \times \left(|\wilson{S}-\wilson{S'}|^2 + |\wilson{P}-\wilson{P'}|^2\right)\,.
\end{aligned}
\end{equation}
is free of tensor couplings at high $q^2$ and would provide access to
$\rho_1^-$ provided scalar couplings vanish.

%
%
%
%--------+---------+---------+---------+---------+---------+---------+---------+
\section{
  Theoretical Inputs
  \label{app:theory:inputs}
}

Here we describe the theoretical treatment of observables and collect numerical
input of the most relevant parameters. The software package {\tt EOS}
\cite{Bobeth:2011gi, Bobeth:2011nj, EOS} is used for the calculation of
observables in $B_s\to \bar\mu\mu$ and $B\to K^{(*)}\bar\ell\ell$ and associated
constraints. Both the likelihood and the prior are defined entirely within EOS.

Concerning numerical input, we refer the reader to~\cite{Beaujean:2013soa} for
the compilation of nuisance parameters and according values, whose uncertainties
are negligible. Further, we adopt the same values for common nuisance parameters of
$i)$ the CKM quark-mixing matrix, $ii)$ the charm and bottom quark masses in the
$\overline{\rm MS}$ scheme, and $iii)$ the parameterisation of subleading corrections
in $1/m_b$ as given in \cite{Beaujean:2013soa}.

The updated prior of the $B_s$ decay constant $f_{B_s}$ enters the branching
ratio of $B_s\to \bar\mu\mu$. We adopt recent updates of the $N_f = (2 + 1)$
FLAG compilation \cite{Aoki:2013ldr}, see \reftab{tab:hadronic:nuisance}, which
averages the results of \cite{Bazavov:2011aa, McNeile:2011ng, Na:2012kp}.
More recent calculations with $N_f = (2+1+1)$ \cite{Dowdall:2013tga} and
$N_f = 2$ \cite{Carrasco:2013zta} are consistent with these averages.

The tensor and scalar amplitudes in $B\to K^{(*)} \bar\ell\ell$ factorize naively,
i.e. they depend only on scalar and tensor $B\to K^{(*)}$ form factors. These amplitudes
are implemented in {\tt EOS} for $B\to K \bar\ell\ell$ and $B\to K^* \bar\ell\ell$ as
given in \cite{Bobeth:2007dw} and \cite{Bobeth:2012vn}, respectively and we refrain from
using respective form factor relations at large and low recoil for the tensor and scalar
$B\to K$ and $B\to K^*$ form factors.

\begin{table}
\begin{center}
\renewcommand{\arraystretch}{1.4}
\begin{tabular}{cccc}
\hline
  Quantity & Prior & Unit & Reference\\
\hline \hline
  \multicolumn{4}{c}{$B_s$ decay constant}
\\
\hline
  $f_{B_s}$   &  $227.7 \pm 4.5$         &  MeV &  \cite{Aoki:2013ldr}\\
\hline \hline
  \multicolumn{4}{c}{$B\to K$ form factors}\\
\hline
  $f_+(0)$    &  $0.34 \pm 0.05$         &  --    &  \cite{Khodjamirian:2010vf, Ball:2004ye}\\
  $f_T(0)$    &  $0.38 \pm 0.06$         &  --    &  \cite{Khodjamirian:2010vf, Ball:2004ye}\\
  $b_1^0$     &  $-4.3^{+0.8}_{-0.9}$    &  --    &  \cite{Khodjamirian:2010vf}\\
  $b_1^+$     &  $-2.1^{+0.9}_{-1.6}$    &  --    &  \cite{Khodjamirian:2010vf}\\
  $b_1^T$     &  $-2.2^{+1.0}_{-2.0}$    &  --    &  \cite{Khodjamirian:2010vf}\\
\hline \hline
\multicolumn{4}{c}{$B\to K^*$ form factors}\\
\hline
\todo{fill in}\\
\hline
\end{tabular}
\renewcommand{\arraystretch}{1.0}
\caption{\label{tab:hadronic:nuisance} Prior distributions of the nuisance parameters
   for hadronic quantities entering inclusive and exclusive decays. Note that
   $f_0(0) = f_+(0)$.
}
\end{center}
\end{table}

As we do not make use of form factor relations for tensor and scalar amplitudes
in $B\to K\bar\ell\ell$, additional nuisance parameters are included for the complete
set of $B\to K$ form factors $f_{+,T,0}$. As a consequence of using the
$z$-parameterization \cite{Khodjamirian:2010vf}, only five nuisance parameters
(listed in \reftab{tab:hadronic:nuisance}) arise due to the kinematic relation
$f_0(0) = f_+(0)$. As prior information, the LCSR results~\cite{Khodjamirian:2010vf}
and \cite{Ball:2004ye} have been averaged (see \reftab{tab:hadronic:nuisance} and
cf. \cite{Beaujean:2013soa}), and supplemented with lattice determinations~\cite{Bouchard:2013eph}.
Contrary to \cite{Beaujean:2013soa}, we do not adapt anymore a log-gamma prior for
the asymmetric uncertainties but rather a continuous, but asymmetric gaussian distribution.
The lattice results are given within a slightly different parameterisation necessitating
a conversion to the parameterisation used here. For this purpose, we generate the
form factors from the parameterisation \cite{Bouchard:2013eph}, including full
correlations, at three values of $q^2 = 17,\, 20,\, 23$ GeV$^2$ finding the results
compiled in \reftab{tab:btok-lattice}. (The $q^2$ values and number of points are chosen
such that the correlation of neighboring points does not exceed $80\%$.)
Subsequently, these ``data'' of the form factors is included in the likelihood
by means of a multivariate Gaussian in addition to experimental measurements.

\begin{table*}
  \begin{center}
    \renewcommand{\arraystretch}{1.3}
    \begin{tabular}{cccc}
        \hline
        $q^2$ [\GeV$^2$] & $17$            & $20$            & $23$\\
        \hline
        $f_0(q^2)$   & $0.62 \pm 0.03$ & $0.72 \pm 0.03$ & $0.87 \pm 0.04$\\
        $f_+(q^2)$   & $1.13 \pm 0.05$ & $1.63 \pm 0.07$ & $2.68 \pm 0.13$\\
        $f_T(q^2)$   & $1.02 \pm 0.06$ & $1.47 \pm 0.08$ & $2.42 \pm 0.18$\\
        \hline
    \end{tabular}\\
    \renewcommand{\arraystretch}{1.0}
    \vspace{2\smallskipamount}
    \renewcommand{\arraystretch}{1.3}
    \begin{tabular}{cccccccccc}
        \hline
        $q^2$ [\GeV$^2$]
             & $17$  & $20$ & $23$ & $17$ & $20$ & $23$ & $17$  & $20$   & $23$ \\
        \hline
   $f_0(17)$ & 1.00  & 0.93 & 0.73 & 0.58 & 0.48 & 0.23 & 0.13  & 0.010 & 0.048 \\
   $f_0(20)$ & --    & 1.00 & 0.87 & 0.41 & 0.45 & 0.24 & 0.059 & 0.093 & 0.075 \\
   $f_0(23)$ & --    & --   & 1.00 & 0.28 & 0.34 & 0.32 & 0.010 & 0.046 & 0.058 \\
   $f_+(17)$ & --    & --   & --   & 1.00 & 0.78 & 0.30 & 0.36  & 0.21  & 0.051 \\
   $f_+(20)$ & --    & --   & --   & --   & 1.00 & 0.71 & 0.22  & 0.26  & 0.20  \\
   $f_+(23)$ & --    & --   & --   & --   & --   & 1.00 &-0.025 & 0.14  & 0.22  \\
   $f_T(17)$ & --    & --   & --   & --   & --   & --   & 1.00  & 0.79  & 0.46  \\
   $f_T(20)$ & --    & --   & --   & --   & --   & --   & --    & 1.00  & 0.85  \\
   $f_T(23)$ & --    & --   & --   & --   & --   & --   & --    & --    & 1.00  \\
        \hline
    \end{tabular}

  \end{center}

    \renewcommand{\arraystretch}{1.0}
    \caption{Mean values, uncertainties (top) and correlation
        information (bottom) of lattice points \cite{Bouchard:2013eph} for the
        $B\to K$ form factors $f_{0,+,T}(q^2)$.
        \label{tab:btok-lattice}
      }
\end{table*}

We change the parameterization of $B\to K^*$ form factors w.r.t. our previous
work \cite{Beaujean:2013soa} slightly and use now the simplified series expansion
(SSE) \cite{Bharucha:2010im}
\begin{align}
  \label{eq:SSE-FF-parameterization}
  F_i(q^2) & =
  \frac{1}{1 - q^2/M_{R,i}^2} \sum_k \alpha_k^i \left[z(q^2) - z(0)\right]^k \,,
\end{align}
with three parameters $\alpha_k^i$ ($k=0,1,2$) per form factor $i = V,\, A_{0,1},
T_{1,23}$ and two ($k=1,2$) for $i= A_{12}, T_2$. The parameters $\alpha_0^i$
correspond to $F_i(q^2 = 0)$ and the kinematic constraints
\begin{align}
  A_{12}(0) & = {\cal F}_{\rm kin}\, A_0(0) \,, &
  T_1(0) & = T_2(0)
\end{align}
are used to eliminate $\alpha_0^{A_{12}, T_2}$, where the kinematic factor
${\cal F}_{\rm kin} \equiv (M_B^2 - M_{K^*}^2)/(8 M_B M_{K^*})$ depends on the
$B$- and $K^*$-meson masses. The residual masses $M_{R,i}$ are set to the values
given in \cite{Straub:2015ica}.

This parameterization allows us to combine consistently available results
of form factors from different non-perturbative approaches, namely LCSR's at
large recoil and lattice at low recoil and implement simultaneously the
kinematic constraints. We determine the parameters $\alpha_i^k$
in a combined fit to form factor predictions of $V, A_{0,1,2},T_{1,2,3}$ from the
LCSR \cite{Straub:2015ica} and of $V, A_{0,1,12},T_{1,2,23}$ from the lattice
\cite{Horgan:2013hoa, Horgan:2015vla} calculations,
including their correlations. Further, we implement the second kinematic constraint
\begin{align}
  \label{eq:A12-constraint-q2max}
  A_{12}(q_{\rm max}^2) & = {\cal F}_{\rm kin}\, A_1(q_{\rm max}^2)
\end{align}
at maximal $q_{\rm max}^2 \equiv (M_B - M_{K^*})^2$. Uninformative flat priors
are chosen for all $\alpha_k^i$
\begin{align}
  \alpha_0^i & \in [0,\, 1] \,, &
  \alpha_{1}^i & \in [-2,\, 2]  \,, &
  \alpha_{2}^i & \in [-3,\, 3] \,.
\end{align}
The results of the LCSR predictions of the form factors have been provided to us
directly by the authors of \cite{Straub:2015ica} at $q^2 = (0.1,\, 4.1,\, 8.1,\,
12.1)\, \mbox{GeV}^2$ for $V,\, A_{0,1,12},\, T_{1,2,3}$, including their $28\times 28$
covariance matrix. \todo{In the case of lattice predictions, we proceed as in our
previous work \cite{Beaujean:2013soa} by generating form factor values of
$V,\, A_{0,1,12},\, T_{1,2,3}$ at $q^2 = (15,\, 19.21)\,\mbox{GeV}^2$ from updated results
\cite{Horgan:2015vla} (arXiv~v2), including their $8\times8$ correlation matrix
see \reftab{tab:btokstar-lattice}.}

\todo{We have compared our choice of three parameters in the SSE \refeq{eq:SSE-FF-parameterization}
versus two parameters only and found that in the latter case lattice form factor
results influence the fit such that form factors tend to be higher than LCSR
predictions at low $q^2$ leading to a worse fit. Hence we prefer the three-parameter
setup as it accommodates better LCSR and lattice results.}

\section{Monte Carlo sampling}
\label{app:mc}
\todo{reference in main text} The marginalization of the posterior is performed
with the package {\tt pypmc}~\cite{beaujean_2015_20045}, which incorporates the
algorithm presented in \cite{Beaujean:2013:PMC, Beaujean:2012uj} and in addition
an implementation of the variational Bayes algorithm. In every analysis we first
run multiple adaptive Markov chains (MCMC) in parallel through
\texttt{pypmc}. If necessary, chains are seeded at the SM point to exclude
solutions in which multiple nuisance parameters---mostly for hadronic
corrections---simultaneously deviate strongly from prior expectations.

In total, there are 19 parameters $\alpha^i_j$ to describe $B\to K^*$ form
factors and most of them are strongly correlated. But it is well known that
strong correlation leads to poor sampling as it can cause the random-walk Markov
chains to spend an excessive amount of  time in regions of low probability and
thus produce spurious peaks. To mitigate this issue, we perform a fit to
form-factor constraints without any experimental data and use the resulting
covariance matrix to transform parameters such that the new parameters are
uncorrelated.

In all but three cases, the Markov chains then give reliable results. But when
we analyze scenarios with $\wilson[NP]{S,P} \ne 0$ and all exp.  constraints,
strong correlations appear again. As a final solution, we then use importance
sampling with the initial proposal function determined by a fit of a Gaussian
mixture to the MCMC samples within the variational Bayes approximation. As the
posterior is unimodal and closely resembles a Gaussian, only a few Gaussian
components are needed; i.e., 3 components proved optimal by the variational
approximation to the model evidence.

In the most challenging run with 62 parameters, we obtain a relative effective
sample size of only 0.038\%. We want to have enough independent samples $N$ such
that the 68\% region is determined with a relative precision of about 1\%. As a
rule of thumb, we consider the ``relative error of the error'' given by
$1/\sqrt{2 N}$~\cite[ch. 37]{Agashe:2014kda}.  In the 62D case, we compute
$1.1\cdot10^6$ importance samples such that $N \approx 3500$ and the estimated
relative error of the error is 1.2\% and thus good enough for our purposes.

To create the smooth marginal plots in Figures
\ref{fig:FHvsCT}--\ref{fig:scalar:SM-EFT}, we apply kernel density estimation
for both MCMC and importance samples using the fast \texttt{figtree}
library~\cite{morariu08figtree}. In the latter case, we additionally crop 500
outliers.

The prior or posterior predictive distribution of an observable $X$ within a
model $M$ in which $X=f(\vecth)$ is a definite function of the
parameters $\vecth$ is given as
\begin{align}
  \label{eq:predictive}
  P(X | M) &= \int \rmdx{\vecth} P(X|\vecth,M) P(\vecth|M)\nn
  &= \int \rmdx{\vecth} \delta(X - f(\vecth)) P(\vecth|M) \,.
\end{align}
We estimate $P(X|M)$ by computing $f(\vecth_i)$ for every sample $\vecth_i \sim
P(\vecth|M)$, then smooth as above.

%
%
%
%--------+---------+---------+---------+---------+---------+---------+---------+
%\section{
%  Standard model predictions
%  \label{app:SM:predictions}
%}

%--------+---------+---------+---------+---------+---------+---------+---------+
%
%  References
%
%--------+---------+---------+---------+---------+---------+---------+---------+

\bibliographystyle{apsrev4-1}
\bibliography{references.bib}

\end{document}